The complex network of trophic interactions in a subAntarctic oceanic Marine Protected

### 2023-04-28 MS_ToSubmit/MS_Burdwood_foodweb.Rmd
Updates Cintia Fraysse

Due to a lack of trophic data resolution for some species inhabiting the study area, we followed the concept of trophic species, here defined as aggregated groups of taxa. In most cases, we followed it when specific data on species, in the taxonomic sense, were not available. In some cases, we collapsed species when taxa shared the same set of predators and prey (trophic similarity, @Martinez1991), one of the aggregation methods that better preserve food web functional properties [@Gauzens2013]. In addition, for endemic species (e.g. bryozoan *Burdwoodipora paguricola*) and other species with no trophic studies so far, we inferred their feeding interactions applying a conservative approach that assumes that the set of prey and predators are at some point preserved in time. In those cases we gathered information from upper taxonomic levels (i.e. Genus, Family, Order, Class, Phylum) as a good proxy variable [@Morales2015; @Pomeranz2018]. Details about this can be found in Supplementary Material (Table S2).

@@

The degree distribution allows to identify important species, such as potential keystone species [@Sole2001; @Dunne2002], generalist/specialist species, and dominant food sources [@Kondoh2010].

We have identified that the majority of the consumers in the food web of the MPA N-BB either has a narrow diet or are specialists, while few present a broad diet. The most evident generalist species are *Patagonotothen guntheri* [@CovattiAle2022], *P. ramsayi* [@Fischer2022], *Dissostichus eleginoides* [@Troccoli2020], *Bathyraja brachyurops* [@Bellegia2008], and *B. griseocauda* [@Bellegia2014], with more than 25 potential prey.
Since these species present mid-trophic positions in the food web (3-4), acting as predator and prey, they might be important links between lower and higher trophic levels. This is in agreement with the sole analysis, using stable isotopes, that exists so far for the trophic structure of the MPA N-BB [@Riccialdelli2020], and resembles other high-latitude marine systems of the Southwest Atlantic and Antarctic regions [@Arkhipkin2013; @Marina2018]. The importance of these particular generalist species also arises since they feed on the benthic and pelagic habitats [@Troccoli2020; @CovattiAle2022; @Fischer2022], linking these realms and contributing to the vertical carbon flow.

On the other hand, few prey are consumed by many predators in the food web of the MPA N-BB. This suggests that there are dominant food sources from which most consumers depend on, and from where the ecosystem energy is being transferred to the upper trophic levels. The most demanded source we identified in this study (i.e. detritus) supports the abundant benthic community of filter-feeders [@Schejter2016]. Furthermore, we found that the second and third-most consumed prey were diatoms and species of *Euphausia*, respectively, which are essential sources for the diverse zooplankton community [@Spinelli2020], mid-TL consumers like the Fuegian sprat *Sprattus fuegensis* [@Padovani2021] and *Patagonotothen ramsayi* [@Fischer2022], and top predators [@RayaRey2022; @Riccialdelli2022].

@@

Describing species’ roles in food webs provides a toolbox with which to assess the significance of species in terms of community's functioning and overall stability [@Thebault2010; @Cirtwill2018]. We used a range of descriptors to characterise the dynamic and multifaceted nature of the species forming the MPA N-BB food web.

Closeness and betweenness are defined as "meso-scale" properties because they consider direct and indirect interactions, therefore describing the focal species’ ability to influence the rest of the species of the food web [@Lai2012].
Closeness quantifies how many steps away species $i$ is from all other species in the food web, and is proportional to how rapidly the indirect effects of the focal species can spread to other species in the network [@Scotti2010]. In the food web of the MPA N-BB, low-TL consumers arise as important in this regard: species of the zooplankton community, *Calanus* and *Euphausia*, and Brachiopoda. Any perturbation affecting these species, such as the recently confirmed contaminants mercury [@Fioramonti2022] and microplastics [@Cossi2021; @DiMauro2022] for the area, should be of concern since it might reach many other species in the food web.
Otherwise, betweenness measures the number of shortest paths between species, providing information on the importance of species as “bridges” for energy transfer: a species with high betweenness takes part in more food chains and therefore affects more energy flows [@Scotti2010]. We have identified the longtail southern cod *Patagonotothen ramsayi* as the most important species in this sense. Moreover, in light of our analysis, other species like the Patagonian toothfish *Dissostichus eleginoides*, the Patagonian cod *Salilota australis*, the yellowfin notothen *Patagonotothen guntheri*, and the Patagonian longfin squid *Doryteuthis gahi* should be considered as relevant in the energy transfer in the ecosystem. All these species have a mid-trophic position in the food web, supporting the 'wasp-waist' control hypothesis for the MPA N-BB [@Riccialdelli2020].

Ecosystems with a pronounced 'wasp-waist' structure are suggested to present a high trophic redundancy, since a large proportion of species would show similar trophic habits [@Cury2000]. The significant negative relationship we found between trophic similarity and trophic level enhances the hypothesis of functional redundancy at low and mid-TL species compared to higher TL species for the MPA N-BB food web [@Riccialdelli2020]. At the same time, our results highlight the uniqueness in terms of trophic role for high-TL predators. Here, not only the expected pelagic animals such as marine mammals and sea birds arise as relevant, but also benthic vertebrate (chondrichthyans *Bathyraja macloviana* and *Squalus acanthias*) and invertebrate species (echinoderms *Diplopteraster clarki* and *Pteraster sp*) are noteworthy. The role that such species play in the MPA N-BB ecosystem is unique and perturbations on them might result in unprecedented changes at the trophic structure and functioning level.

Species' role can be also assessed in a module-based context. Among the varying numbers of topological roles in which species can be divided, two are remarkable: 'module connector' and 'network connector'. Here again the longtail southern cod *Patagonotothen ramsayi* is emphasized as an influential species for the connectivity of the food web, which agrees with the suggested 'wasp-waist' control for the MPA N-BB ecosystem [@Riccialdelli2020]. But most importantly, our results broaden this issue by pointing out that there are other species, belonging to a wide range of trophic positions (1 to 3.86) and representing 20 different functional groups, that should be considered. Thus, we propose that it is the diversity of species (benthic and pelagic) what maintains the connectivity of the food web, therefore contributing to the trophic structure and ecosystem's stability.

@@

Some species of sessile suspension feeders in high-latitude marine ecosystems, such as sponges, ascidians and octocorals, avoid predation by producing secondary metabolites that function as a chemical defense [@NunezPons2010; @Moles2015; @Prieto2022]. Although this was not reported at the MPA N-BB, there are few studies that reported it other locations in species that were found in the MPA N-BB [@RojoDeAlmeida2010]. Thus, the interactions that we considered between these species and their predators should be taken with caution.

@@

Spatial variation of study area: Banco Burdwood west-east gradient.
In the oceanic BB area, complex oceanographic conditions generate an internal spatial heterogeneity, mainly along its longitudinal axis [@Matano2019], and consequently plankton communities respond with different spatial distributions [@Spinelli2020].
@Kortsch2019

@@

* Incorporate habitat heterogeneity, since benthic community in the BB and deep flanks is different [@Schejter2021]. El caso de las ascidias (Tatián)

This might lead to distinct food web properties, hence differences in ecosystem functioning and response to anthropogenic and environmental perturbations.
@Cordone2020

@@

We compiled information on the species and trophic diversity of the Marine Protected Area Namuncurá - Banco Burdwood, generating an unprecedented, well-resolved network of trophic interactions for a sub-Antarctic ecosystem, identifying the complexity and structure of the system, and the main species role in a network framework. Particular properties at the network level allowed us to identify ecosystem's vulnerability and potential response to perturbations in the presence of highly-connected species, with a rearrange structure maintaining their functions due to its potential resilience and resistance.

We identified several species as important in terms of different aspects of trophic structure and functioning, and response to perturbations (i.e. environmental/anthropogenic changes). On one hand, we suggest that generalist species, mostly fishes, play a crucial role in the ecosystem's bentho-pelagic coupling process. At the same time, we propose that other species besides the longtail southern cod *Patagonotothen ramsayi* and the Fueguian sprat *Sprattus fuegensis* should be considered as relevant energy transfers for the ecosystem. Finally, we argue that it is the diversity of species, representing the benthic and pelagic habitats, what maintains the connectivity of the food web against perturbations, therefore contributing to the structure and stability of the ecosystem.

>>>>

Due to a lack of trophic data resolution for some species inhabiting the study area, we followed the concept of trophic species, here defined as aggregated groups of taxa. In most cases, we followed it when specific data on species, in the taxonomic sense, were not available. In some cases, we collapsed species when taxa shared the same set of predators and prey (trophic similarity, @Martinez1991), one of the aggregation methods that better preserve food web functional properties [@Gauzens2013]. In addition, for endemic species (e.g. bryozoan *Burdwoodipora paguricola*) and other species with no trophic studies so far, we inferred their feeding interactions applying a conservative approach that assumes that the set of prey and predators are at some point preserved in time. In those cases we gathered information from upper taxonomic levels (i.e. Genus, Family, Order, Class, Phylum) as a good proxy variable [@Morales-Castilla2015; @Pomeranz2019]. Details about this can be found in Supplementary Material (Table S2).

@@

The degree distribution allows identifying important species, such as potential keystone species [@Sole2001; @Dunne2002], generalist/specialist species, and dominant food sources [@Kondoh2010].

We have identified that most of the consumers in the food web of the MPA N-BB either has a narrow diet or are specialists, while few present a broad diet. The most evident generalist species are *Patagonotothen guntheri* [@CovattiAle2022], *P. ramsayi* [@Fischer2022], *Dissostichus eleginoides* [@Troccoli2020], *Bathyraja brachyurops* [@Belleggia2008a], and *B. griseocauda* [@Bellegia2014], with more than 25 potential prey.
Since these species present mid-trophic positions in the food web (3-4), acting as predator and prey, they might be important links between lower and higher trophic levels. This result is in agreement with the sole analysis, using stable isotopes, that exists so far for the trophic structure of the MPA N-BB [@Riccialdelli2020], and resembles other high-latitude marine systems of the Southwest Atlantic and Antarctic regions [@Arkhipkin2013; @Marina2018]. The importance of these particular generalist species also arises since they feed on the benthic and pelagic habitats [@Troccoli2020; @CovattiAle2022; @Fischer2022], linking these realms and contributing to the vertical carbon flow.

On the other hand, a low number of prey are consumed by many predators in the food web of the MPA N-BB. This suggests that there are dominant food sources from which most consumers depend on and from where the ecosystem energy is being transferred to the upper trophic levels. The most demanded source we identified in this study (i.e. detritus) supports the abundant benthic community of filter-feeders [@Schejter2016]. Furthermore, we found that the second and third-most consumed prey were diatoms and species of *Euphausia*, respectively, which are essential sources for the diverse zooplankton community [@Spinelli2020], mid-TL consumers like the Fuegian sprat *Sprattus fuegensis* [@Padovani2021] and *Patagonotothen ramsayi* [@Fischer2022], and top predators [@RayaRey2022; @Riccialdelli2022].

@@

Describing species’ roles in food webs provides a toolbox to assess the significance of species in terms of community's functioning and overall stability [@Thebault2010; @Cirtwill2018]. We used a range of descriptors to characterise the dynamic and multifaceted nature of the species forming the MPA N-BB food web.

Closeness and betweenness are defined as "mesoscale" properties because they consider direct and indirect interactions, therefore describing the focal species’ ability to influence the rest of the species of the food web [@Lai2012].
Closeness quantifies how many steps away species $i$ is from all other species in the food web, and is proportional to how rapidly the indirect effects of the focal species can spread to other species in the network [@Scotti2010]. In the food web of the MPA N-BB, low-TL consumers arise as important in this regard: species of the zooplankton community, *Calanus* and *Euphausia*, and Brachiopoda. Any perturbation affecting these species, such as the recently confirmed contaminants mercury [@Fioramonti2022] and microplastics [@Cossi2021; @DiMauro2022] for the area, should be of concern since it might reach many other species in the food web.
Otherwise, betweenness measures the number of shortest paths between species, providing information on the importance of species as “bridges” for energy transfer: a species with high betweenness takes part in more food chains and therefore affects more energy flows [@Scotti2010]. We have identified the longtail southern cod *Patagonotothen ramsayi* as the most important species in this sense. Moreover, in light of our analysis, other species like the Patagonian toothfish *Dissostichus eleginoides*, the Patagonian cod *Salilota australis*, the yellowfin notothen *Patagonotothen guntheri*, and the Patagonian longfin squid *Doryteuthis gahi* should be considered as relevant in the energy transfer in the ecosystem. All these species have a mid-trophic position in the food web, supporting the 'wasp-waist' control hypothesis for the MPA N-BB [@Riccialdelli2020].

Ecosystems with a pronounced 'wasp-waist' structure are suggested to present a high trophic redundancy, since many species would show similar trophic habits [@Cury2000]. The significant negative relationship between trophic similarity and trophic level enhances the hypothesis of functional redundancy at low and mid-TL species compared to higher TL species for the MPA N-BB food web [@Riccialdelli2020]. At the same time, our results highlight the uniqueness in terms of the trophic role of high-TL predators. Here, not only the expected pelagic animals such as marine mammals and sea birds arise as relevant, but also benthic vertebrate (chondrichthyans *Bathyraja macloviana* and *Squalus acanthias*) and invertebrate species (echinoderms *Diplopteraster clarki* and *Pteraster sp*) are noteworthy. The role that such species play in the MPA N-BB ecosystem is unique and perturbations on them might result in unprecedented changes at the trophic structure and functioning level.

Species' role can also be assessed in a module-based context. Among the varying numbers of topological roles in which species can be divided, two are remarkable: 'module connector' and 'network connector'. Here again, the longtail southern cod *Patagonotothen ramsayi* is emphasized as an influential species for the connectivity of the food web, which agrees with the suggested 'wasp-waist' control for the MPA N-BB ecosystem [@Riccialdelli2020]. But most importantly, our results broaden this issue by pointing out that there are other species, belonging to a wide range of trophic positions (1 to 3.86) and representing 20 different functional groups, that should be considered. Thus, we propose that the diversity of species (benthic and pelagic) maintains the connectivity of the food web, therefore contributing to the trophic structure and ecosystem's stability.

@@

Some species of sessile suspension feeders in high-latitude marine ecosystems, such as sponges, ascidians and octocorals, avoid predation by producing secondary metabolites that function as a chemical defense [@Nunez-Pons2010; @Moles2015; @Prieto2022]. Although this was not reported at the MPA N-BB, there are few studies that reported it other locations in species that were found in the MPA N-BB [@RojodeAlmeida2010]. Thus, the interactions that we considered between these species and their predators should be taken with caution.

@@

Spatial variation of study area: Banco Burdwood west-east gradient.
In the oceanic BB area, complex oceanographic conditions generate an internal spatial heterogeneity, mainly along its longitudinal axis [@Matano2019], and consequently, plankton communities respond with different spatial distributions [@Spinelli2020].
@Kortsch2019

@@

* Incorporate habitat heterogeneity, since the benthic community in the BB and deep flanks differs [@Schejter2021]. This heterogeneity might lead to distinct food web properties, hence differences in ecosystem functioning and response to anthropogenic and environmental perturbations [@Cordone2020].
El caso de las ascidias (Tatián)

@@

We compiled information on the species and trophic diversity of the Marine Protected Area Namuncurá - Banco Burdwood, generating an unprecedented, well-resolved network of trophic interactions for a sub-Antarctic ecosystem, identifying the complexity and structure of the system, and the main species role in a network framework. Particular properties at the network level allowed us to identify the ecosystem's vulnerability and potential response to perturbations in the presence of highly-connected species, with a rearranged structure maintaining their functions due to its potential resilience and resistance.

We identified several species as important regarding different aspects of trophic structure and functioning, and response to perturbations (i.e. environmental/anthropogenic changes). On the one hand, we suggest that generalist species, mainly fishes, play a crucial role in the ecosystem's bentho-pelagic coupling process. At the same time, we propose that other species besides the longtail southern cod *Patagonotothen ramsayi* and the Fueguian sprat *Sprattus fuegensis* should be considered relevant energy transfers for the ecosystem. Finally, we argue that it is the diversity of species, representing the benthic and pelagic habitats, that maintains the connectivity of the food web against perturbations, therefore contributing to the structure and stability of the ecosystem.
